---
title: "Actvity Monitoring Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(lubridate)
```

#1) Code for reading in the dataset and/or processing the data

```{r}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
activity <- read.csv('activity.csv')

activity$date <- ymd(activity$date)

activity$weekend <- as.factor(ifelse(weekdays(activity$date)=="Saturday" | weekdays(activity$date)=="Sunday","weekend","weekday"))

activity$dayofweek <- as.factor(weekdays(activity$date))

head(activity)
```

#2)Histogram of the total number of steps taken each day

```{r}
databydate <- activity %>% 
  select(date, steps) %>% 
  group_by(date) %>% 
  summarize(tsteps= sum(steps)) %>%
  na.omit()


hist(databydate$tsteps, xlab = "Total daily Steps",main="Histogram of Total Steps by day", breaks = 20)

```

#3) Mean and median number of steps taken each day

```{r}
daily_steps<- activity %>%
  select(date,steps) %>%
  group_by(date) %>%
  summarize(mean_steps_daily = mean(steps),median_steps_daily = median(steps)) %>%
  na.omit()

daily_steps

plot(daily_steps$date,daily_steps$mean_steps_daily,type="l")

```

#4) Time series plot of the average number of steps taken

```{r}

steps_by_interval <- activity%>% 
  select(interval, steps) %>% 
  na.omit() %>% 
  group_by(interval) %>% 
  summarize(total_steps= mean(steps)) 

ggplot(steps_by_interval, aes(x=interval, y=total_steps))+ geom_line(col="red")

```

#5) The 5-minute interval that, on average, contains the maximum number of steps

```{r}
steps_by_interval <- activity%>% 
  select(interval, steps) %>% 
  na.omit() %>% 
  group_by(interval) %>% 
  summarize(total_steps= mean(steps))

steps_by_interval[which(steps_by_interval$total_steps == max(steps_by_interval$total_steps)),]

```

#6) Code to describe and show a strategy for imputing missing data

```{r}
## 1. Code to describe and show a strategy for imputing missing data
head(activity)

missingVals <- sum(is.na(activity$steps))

## 2. Devise a strategy for filling in all of the missing values in the dataset.

replace_with_mean <- function(x) {
  
 replace(x, is.na(x), mean(x, na.rm = TRUE))
}

meandata <- activity%>% group_by(interval) %>% mutate(steps= round(replace_with_mean(steps),2))

head(meandata)

```

#7) Histogram of the total number of steps taken each day after missing values are imputed

```{r}
daily_steps_after_imputing <- meandata %>% 
  select(date, steps) %>% 
  group_by(date) %>% 
  summarize(total_steps= sum(steps)) %>%
  na.omit()

hist(daily_steps_after_imputing$total_steps, xlab = "Total daily Steps",main="Histogram of Total Steps by day", breaks = 20,ylim = c(0,20),col = "sky blue")

```

#8) Panel plot comparing the average number of steps taken per 5-minute interval 
#across weekdays and weekends

```{r}
meansteps <- meandata %>% group_by(interval,weekend) %>%  summarise(average = mean(steps))

qplot(interval,average,data=meansteps,geom="line",facets=weekend~.,col=weekend,xlab="5-minute interval",ylab="average number of steps",main="Average steps pattern between Weekday and Weekend")

ggplot(meansteps,aes(x=interval, y=average, color=weekend)) + geom_line()+
  facet_grid(weekend ~.) + xlab("Interval") + ylab("Mean of Steps") +
  ggtitle("Comparison of Average Number of Steps in Each Interval")

```

